---
title: "Palmer's Penguins"
format: html
editor: visual
---

```{r}
#| message: false
library(tidyverse)
library(palmerpenguins)
```

```{r}
colnames(penguins)

# Colorblind safe color theme
okabe_ito <- c("#E69F00", "#56B4E9", "#009E73", "#F0E442", 
               "#0072B2", "#D55E00", "#CC79A7", "#000000")
```

## Distribution of a single trait

### Bill length

```{r}
ggplot(data = penguins) +
  geom_histogram(mapping = aes(x = bill_length_mm), bins = 10) +
  facet_wrap(~species) +
  theme_linedraw()
```

Adelie have the smallest bill length with Chinstrap and Gentoo have similar distributions

## Relationship between multiple traits

### Do the same species penguins on different islands have the same bill length?

```{r}
ggplot(data = penguins) +
  geom_boxplot(mapping = aes(x = island, y = bill_length_mm)) +
  geom_jitter(mapping = aes(x = island, y = bill_length_mm)) +
  facet_wrap(~species) +
  theme_linedraw()
```

Only Adelie penguins are found on the all three islands and their bill length do not seem to differ. Chinstrap and Gentoo are only found on 1 island.

### Does body mass impact bill length?

```{r}
ggplot(data = penguins) +
  geom_point(mapping = aes(x = body_mass_g, y = bill_length_mm)) +
  geom_smooth(mapping = aes(x = body_mass_g, y = bill_length_mm)) +
  theme_linedraw()
```

Across all penguins, the trend looks the same but there is a group of individuals with a higher bill length relative to body mass. Could that be due to a species?

```{r}
ggplot(data = penguins) +
  geom_point(mapping = aes(x = body_mass_g, y = bill_length_mm, color = species)) +
  geom_smooth(mapping = aes(x = body_mass_g, y = bill_length_mm, color = species)) +
  scale_color_manual(values = okabe_ito) +
  theme_linedraw()
```

Yes. That higher bill_length ratio was actually a different penguin species, Chinstrap. These curve lines look strange though. Can I improve them by using a linear model instead of the default?

```{r}
ggplot(data = penguins) +
  geom_point(mapping = aes(x = body_mass_g, y = bill_length_mm, color = species)) +
  geom_smooth(
    mapping = aes(x = body_mass_g, y = bill_length_mm, color = species),
    method = lm
  ) +
  scale_color_manual(values = okabe_ito) +
  theme_linedraw()
```

Yes. That looks nicer.

## Levels of Sexual Dimorphism

### What is the sex balance of each species?

```{r}
ggplot(data = penguins) +
  geom_bar(mapping = aes(x = species, fill = sex)) +
  scale_fill_manual(values = okabe_ito) +
  theme_linedraw()
```

Looks even to me. I wish I could show the data as percentages but not sure how. Let's look at this plot from the worksheet

### Example plot from datasheet

```{r}
penguins %>%
  ggplot(mapping = aes(x = flipper_length_mm, y = sex, color = sex)) +
  geom_boxplot(outlier.alpha = 0, alpha = 0) +
  geom_jitter(width = 0, size = 0.1) +
  coord_flip() +
  facet_wrap( ~ species)
```

Males have larger flipper lengths in all three species. Some recorded penguins don't have sex data.
